---
title: "Advanced R programming final assignment"
author: "Zhandi Liao"
date: "April 29, 2018"
output: html_document
---
Coursera "Adanced R programming" Functional and Object-Oriented Programming
==================================================================================
## Part 1: Factorial Function-- code

The objective of Part 1 is to write a function that computes the factorial of an integer greater than or equal to 0. Recall that the factorial of a number n is n * (n-1) * (n - 2) * . * 1. The factorial of 0 is defined to be 1. Before taking on this part of the assignment, you may want to review the section on Functional Programming in this course.

For this Part you will need to write four different versions of the Factorial function:

Factorial_loop: a version that computes the factorial of an integer using looping (such as a for loop)
Factorial_reduce: a version that computes the factorial using the reduce() function in the purrr package. Alternatively, you can use the Reduce() function in the base package.
Factorial_func: a version that uses recursion to compute the factorial.
Factorial_mem: a version that uses memoization to compute the factorial.

```{r part1_code, echo=TRUE}
library(purrr)
library(microbenchmark)
library(magrittr)
library(tidyr)
library(dplyr)

factorial_loop <- function(n) {
  stopifnot(n>=0)
  ans = 1
  if(n==0){
      ans
  } else for(i in 1:n){
      ans <- ans*i
  }
  return(ans)
}

factorial_reduce <- function(n){
  stopifnot(n>=0)
  if(n==0){return(1)}
  else return(reduce(1:n,function(x,y){as.numeric(x*y)}))
}

factorial_func <- function(n){
  stopifnot(n>=0)
  if(n==0){return(1)}
  else return(as.numeric(n*factorial_func(n-1)))
}

fac_table <- c(1,1,rep(NA,500))

factorial_mem <- function(n){
  stopifnot(n>=0)
  if(!is.na(fac_table[n+1])){
    return(fac_table[n+1])
  } else {
    fac_table[n] <<- factorial_mem(n-1)
    return(as.numeric(n*fac_table[n]))
  }
}
```

## Part 1: Factorial Function-- output

After writing your four versions of the Factorial function, use the microbenchmark package to time the operation of these functions and provide a summary of their performance. In addition to timing your functions for specific inputs, make sure to show a range of inputs in order to demonstrate the timing of each function for larger inputs.

```{r part1_output,echo= TRUE}
loop_data <- map(1:100,function(x){microbenchmark(factorial_loop(x),times=101)$time})
loop_data <- as.data.frame(loop_data)
names(loop_data) <- 1:100
loop_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)

reduce_data <- map(1:100,function(x){microbenchmark(factorial_reduce(x),times = 101)$time})
reduce_data <- as.data.frame(reduce_data)
names(reduce_data) <- 1:100
reduce_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)

func_data <- map(1:100,function(x){microbenchmark(factorial_func(x),times = 101)$time})
func_data <- as.data.frame(func_data)
names(func_data) <- 1:100
func_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)

mem_data <- map(1:100,function(x){microbenchmark(factorial_mem(x),times = 101)$time})
mem_data <- as.data.frame(mem_data)
names(mem_data) <- 1:100
mem_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)
```

Generate the export data with the mean running time
```{r part1_output, echo=TRUE}
mydata<-data.frame(loop = loop_data$mean,
                   reduce = reduce_data$mean,
                   func = func_data$mean,
                   mem = mem_data$mean,
                   row.names = loop_data$num)
write.csv(mydata, file = "part1_output.txt")
```

