---
title: "Advanced R programming final assignment"
author: "Zhandi Liao"
date: "April 29, 2018"
output: html_document
---
Coursera "Adanced R programming" Functional and Object-Oriented Programming
==================================================================================
## Part 1: Factorial Function-- code

The objective of Part 1 is to write a function that computes the factorial of an integer greater than or equal to 0. Recall that the factorial of a number n is n * (n-1) * (n - 2) * . * 1. The factorial of 0 is defined to be 1. Before taking on this part of the assignment, you may want to review the section on Functional Programming in this course.

For this Part you will need to write four different versions of the Factorial function:

Factorial_loop: a version that computes the factorial of an integer using looping (such as a for loop)
Factorial_reduce: a version that computes the factorial using the reduce() function in the purrr package. Alternatively, you can use the Reduce() function in the base package.
Factorial_func: a version that uses recursion to compute the factorial.
Factorial_mem: a version that uses memoization to compute the factorial.

```{r part1_code, echo=TRUE}
library(purrr)
library(microbenchmark)
library(magrittr)
library(tidyr)
library(dplyr)

factorial_loop <- function(n) {
  stopifnot(n>=0)
  ans = 1
  if(n==0){
      ans
  } else for(i in 1:n){
      ans <- ans*i
  }
  return(ans)
}

factorial_reduce <- function(n){
  stopifnot(n>=0)
  if(n==0){return(1)}
  else return(reduce(1:n,function(x,y){as.numeric(x*y)}))
}

factorial_func <- function(n){
  stopifnot(n>=0)
  if(n==0){return(1)}
  else return(as.numeric(n*factorial_func(n-1)))
}

fac_table <- c(1,1,rep(NA,500))

factorial_mem <- function(n){
  stopifnot(n>=0)
  if(!is.na(fac_table[n+1])){
    return(fac_table[n+1])
  } else {
    fac_table[n] <<- factorial_mem(n-1)
    return(as.numeric(n*fac_table[n]))
  }
}
```

## Part 1: Factorial Function-- output

After writing your four versions of the Factorial function, use the microbenchmark package to time the operation of these functions and provide a summary of their performance. In addition to timing your functions for specific inputs, make sure to show a range of inputs in order to demonstrate the timing of each function for larger inputs.

```{r part1_output,echo= TRUE}
loop_data <- map(1:100,function(x){microbenchmark(factorial_loop(x),times=101)$time})
loop_data <- as.data.frame(loop_data)
names(loop_data) <- 1:100
loop_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)

reduce_data <- map(1:100,function(x){microbenchmark(factorial_reduce(x),times = 101)$time})
reduce_data <- as.data.frame(reduce_data)
names(reduce_data) <- 1:100
reduce_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)

func_data <- map(1:100,function(x){microbenchmark(factorial_func(x),times = 101)$time})
func_data <- as.data.frame(func_data)
names(func_data) <- 1:100
func_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)

mem_data <- map(1:100,function(x){microbenchmark(factorial_mem(x),times = 101)$time})
mem_data <- as.data.frame(mem_data)
names(mem_data) <- 1:100
mem_data %<>%
  gather(num,time) %>%
  group_by(num) %>%
  summarise(min = min(time),
            median = median(time),
            max=max(time),
            mean=mean(time),
            std=sd(time)) %>%
  mutate(num = as.numeric(num)) %>%
  arrange(num)
```

Generate the export data with the mean running time
```{r part1_output, echo=TRUE}
mydata<-data.frame(loop = loop_data$mean,
                   reduce = reduce_data$mean,
                   func = func_data$mean,
                   mem = mem_data$mean,
                   row.names = loop_data$num)
write.csv(mydata, file = "part1_output.txt")
```


## Part 2: Longitudinal Data Class and Methods

The purpose of this part is to create a new class for representing longitudinal data, which is data that is collected over time on a given subject/person. This data may be collected at multiple visits, in multiple locations. You will need to write a series of generics and methods for interacting with this kind of data.

The data for this part come from a small study on indoor air pollution on 10 subjects. Each subject was visited 3 times for data collection. Indoor air pollution was measured using a high-resolution monitor which records pollutant levels every 5 minutes and the monitor was placed in the home for about 1 week. In addition to measuring pollutant levels in the bedroom, a separate monitor was usually placed in another room in the house at roughly the same time.

The variables in the dataset are

id: the subject identification number
visit: the visit number which can be 0, 1, or 2
room: the room in which the monitor was placed
value: the level of pollution in micrograms per cubic meter
timepoint: the time point of the monitor value for a given visit/room
You will need to design a class called "LongitudinalData" that characterizes the structure of this longitudinal dataset. You will also need to design classes to represent the concept of a "subject", a "visit", and a "room".

In addition you will need to implement the following functions

make_LD: a function that converts a data frame into a "LongitudinalData" object
subject: a generic function for extracting subject-specific information
visit: a generic function for extracting visit-specific information
room: a generic function for extracting room-specific information
For each generic/class combination you will need to implement a method, although not all combinations are necessary (see below). You will also need to write print and summary methods for some classes (again, see below).

To complete this Part, you can use either the S3 system, the S4 system, or the reference class system to implement the necessary functions. It is probably not wise to mix any of the systems together, but you should be able to compete the assignment using any of the three systems. The amount of work required should be the same when using any of the systems.

```{r part2,echo=TRUE}
## Read in the data
library(readr)
library(magrittr)
source("oop_code.R")
## Load any other packages that you may need to execute your code

data <- read_csv("data/MIE.csv")
x <- make_LD(data)
print(class(x))
print(x)

## Subject 10 doesn't exist
out <- subject(x, 10)
print(out)

out <- subject(x, 14)
print(out)

out <- subject(x, 54) %>% summary
print(out)

out <- subject(x, 14) %>% summary
print(out)

out <- subject(x, 44) %>% visit(0) %>% room("bedroom")
print(out)

## Show a summary of the pollutant values
out <- subject(x, 44) %>% visit(0) %>% room("bedroom") %>% summary
print(out)

out <- subject(x, 44) %>% visit(1) %>% room("living room") %>% summary
print(out)

```
