---
title: "R_prog_envir_week_4"
author: "Zhandi Liao"
date: "March 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

daily_SPEC_2014.csv.bz2: a compressed CSV file containing daily measurements of particulate matter chemical constituents in the United States for the year 2014. Note that you should NOT have to decompress this file. The data are measured at a network of federal, state, and local monitors and assembled by the EPA. In this dataset, the "Arithmetic Mean" column provides the level of the indicated chemical constituent and the "Parameter.Name" column provides the name of the chemical constituent. The combination of a "State Code", a "County Code", and a "Site Num", uniquely identifies a monitoring site (the location of which is provided by the "Latitude" and "Longitude" columns).

aqs_sites.xlsx: An excel spreadsheet containing metadata about each of the monitoring sites in the United States where pollution measurements are made. In particular, the "Land Use" and "Location Setting" variables contain information about what kinds of areas the monitors are located in (i.e. "residential" vs. "forest").

For this project, the following packages are required: "dplyr","tidyr","readr","readxl"

```{r loading}
library(tidyverse)
library(readxl)

file_url <- "https://d18ky98rnyall9.cloudfront.net/_f018d9fe5547b1a722ce260af0fa71af_quiz_data.zip?Expires=1522281600&Signature=c5VtzeNXD~Jnr8hzLlXGu8OglCnBUln7fUeWoUJFr~A0HnDgRCTKuETaqQ38Rwt8FH0q31KZ0jcNIf2JZj1Sldddh0QswQz35tsAMDrLABb8WAxwAmvgvCRd~znyEUMcAm51PPnoNHySEQLB6MO7mzgHvCTqtihbs5rB669iy~g_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A"
```

## Getting the files

Files are fetched from the online link

```{r getting file,cache=TRUE}
if(!file.exists("~/R_envir_week4")) {
  dir.create("~/R_envir_week4")
}
# check if file has been downloaded then download the file
if(!file.exists("~/R_envir_week4/quizdata.zip")) {
datafile <- download.file(file_url,destfile = "~/R_envir_week4/quizdata.zip", mode = "wb")}
unzip("~/R_envir_week4/quizdata.zip",exdir = "~/R_envir_week4")
filepath<-"~/R_envir_week4/data"
sites <- read_xlsx(paste0(filepath,"/aqs_sites.xlsx"))
daily_spec <- read_csv(paste0(filepath,"/daily_SPEC_2014.csv.bz2"))
names(daily_spec)<-gsub(" ",".",names(daily_spec))
names(sites)<-gsub(" ",".",names(sites))
```

## Question 1.
What is the average Arithmetic.Mean for "Bromine PM2.5 LC" in the state of Wisconsin in the dataset?

```{r Q1}
a1<- daily_spec %>%
    filter(Parameter.Name == "Bromine PM2.5 LC" & 
             State.Name == "Wisconsin") %>%
    select(Arithmetic.Mean)
mean(a1$Arithmetic.Mean,na.rm = TRUE)
```

## Question 2.
Calculate the average of each chemical constituent across all states, monitoring sites and all time points.

Which constituent Parameter.Name has the highest average level?
```{r Q2}
a2<-daily_spec %>%
    group_by(factor(Parameter.Name)) %>%
    summarize(mean = mean(Arithmetic.Mean,na.rm = TRUE)) %>%
    arrange(desc(mean))
a2[a2[,1] == "Sulfur PM2.5 LC"|
   a2[,1] == "Sodium PM2.5 LC"|
   a2[,1] == "OC CSN Unadjusted PM2.5 LC TOT"|
   a2[,1] == "EC2 PM2.5 LC",]
```

## Question 3.
Which monitoring site has the highest average level of "Sulfate PM2.5 LC" across all time?

Indicate the state code, county code, and site number.
```{r Q3}
a3<-daily_spec %>%
    filter(Parameter.Name == "Sulfate PM2.5 LC" & !is.na(Arithmetic.Mean)) %>%
    arrange(desc(Arithmetic.Mean)) %>%
    select(State.Code,County.Code,Site.Num,Arithmetic.Mean) %>%
    slice(1:5)
print(a3)
```

## Question 4.
What is the absolute difference in the average levels of "EC PM2.5 LC TOR" between the states California and Arizona, across all time and all monitoring sites?
```{r Q4}
a4<-daily_spec %>%
    filter(Parameter.Name == "EC PM2.5 LC TOR" &
             (State.Name == "California" | State.Name == "Arizona")) %>%
    group_by(factor(State.Name)) %>%
    summarize(mean = mean(Arithmetic.Mean,na.rm = TRUE))
abs(a4[1,2]-a4[2,2])
```

## Question 5.
What is the median level of "OC PM2.5 LC TOR" in the western United States, across all time? Define western as any monitoring location that has a Longitude LESS THAN -100.
```{r Q5}
a5<-daily_spec %>%
    filter(Longitude < -100 & Parameter.Name == "OC PM2.5 LC TOR") %>%
    select(Arithmetic.Mean) %>%
    summarize(median = median(Arithmetic.Mean))
print(a5)
```

## Question 6.
Use the readxl package to read the file aqs_sites.xlsx into R (you may need to install the package first). This file contains metadata about each of the monitoring sites in the EPA's monitoring system. In particular, the "Land Use" and "Location Setting" variables contain information about what kinds of areas the monitors are located in (i.e. "residential" vs. "forest").

How many monitoring sites are labelled as both RESIDENTIAL for "Land Use" and SUBURBAN for "Location Setting"?
```{r Q6}
a6<-sites %>%
    filter(Land.Use == "RESIDENTIAL" & Location.Setting == "SUBURBAN") %>%
    summarize(N = n())
print(a6)
```

## Question 7.
What is the median level of "EC PM2.5 LC TOR" amongst monitoring sites that are labelled as both "RESIDENTIAL" and "SUBURBAN" in the eastern U.S., where eastern is defined as Longitude greater than or equal to -100?
a7<-sites %>%
    filter(Land.Use == "RESIDENTIAL" & Location.Setting == "SUBURBAN") %>%